---
title: "NSW Case Ascertainment"
description: |
  Using the apyramid package to compare how Covid-19 cases are discovered by age group
author:
  - name: Sean De Boo
    url: https://seandeboo.github.io/hds-blog/
date: 2022-06-28
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
library(tidyverse)
library(apyramid)

knitr::opts_chunk$set()
```


``` {r gather, echo=FALSE}

NSW <- read_csv("https://data.nsw.gov.au/data/dataset/3dc5dc39-40b4-4ee9-8ec6-2d862a916dcf/resource/4b03bc25-ab4b-46c0-bb3e-0c839c9915c5/download/confirmed_cases_table2_age_group_agg.csv")

## make age_group a factor
NSW$age_group <- as.factor(NSW$age_group)

```

## NSW Case data

### 1. NSW cases in the past 2 weeks

Looking at case data for the past 2 weeks, a large portion are in the 0-19 age group. However, the 70+ age group, which is the main user of PCR tests, also had a significant number of cases

``` {r plot, fig.width = 10}

NSWc <- NSW %>%
  filter(notification_date >= '2022-06-12') %>% 
  group_by(age_group, confirmed_by_pcr) %>% 
  summarise(PCR = sum(confirmed_cases_count))


age_pyramid(data = NSWc, age_group = 'age_group',
            split_by = 'confirmed_by_pcr', count = 'PCR')


```

### 2. NSW cases since RAT tests became available

Looking back to the start of Rapid Antigen Tests (RAT) being sold in NSW, we can see the pattern has been roughly that same. A large amount of the cases are aged 0-19 years old and the point where PCR becomes the main test type is around age 50 and over.

``` {r since_recorded, fig.width=10, echo=FALSE}
NSWd <- NSW %>%
  filter(notification_date >= '2022-01-20') %>% 
  group_by(age_group, confirmed_by_pcr) %>% 
  summarise(PCR = sum(confirmed_cases_count))


age_pyramid(data = NSWd, age_group = 'age_group',
            split_by = 'confirmed_by_pcr', count = 'PCR')


```

Data can be access from [NSW COVID-19 cases by age range](https://tinyurl.com/4ajzzxws)