---
title: "Queensland Covid Cases in 2022"
description: |
  A look at the daily covid case counts in Qld so far in 2022.
author:
  - name: Sean De Boo
    url: https://seandeboo.github.io/hds-blog/
date: 06/11/2022
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)

```

Covid case data by 5 year cohort is found at the Qld data portal [Qld Data Portal](https://www.data.qld.gov.au/dataset/queensland-covid-19-case-line-list-age-groups).

The data is imported as a csv file with the notifcation date of the case as well as the 5 year age
bracket the case falls into.

```{r gather and transform, warning=FALSE}
qld <- read_csv('https://www.data.qld.gov.au/dataset/5efaa096-4480-4540-88be-a10ababd9f49/resource/a14317b7-2fca-41b7-8294-9a1f7a085b0f/download/opendata_qld_covidcase_agegrp.csv')

qld$AGE_GROUP_5Y <- as.factor(qld$AGE_GROUP_5Y)
qld$NOTIFICATION_DATE <- as.Date(qld$NOTIFICATION_DATE, format = "%d/%m/%Y")

qld

```

The date is imported as a character and was converted using as.Date(). Also the Age Groups were converted to a factor for easier grouping.

The next step is to filter data since 1st of December in 2021 which was when the border was opened, and local cases were allowed to grow.

```{r transform, echo=TRUE}
qld2 <- qld %>%
  filter(NOTIFICATION_DATE > '2021-12-01') %>%
  filter(AGE_GROUP_5Y != '100+ years') %>%
  filter(!is.na(AGE_GROUP_5Y)) %>% 
  group_by(NOTIFICATION_DATE) %>% 
  count(AGE_GROUP_5Y)
```

removed the plots for NA and 100+ age groups


``` {r plot, fig.width=12, fig.height=8}
ggplot() +
ggtitle('QLD Confirmed cases')+ 
xlab('Date') + ylab('Number of Cases')+ 
geom_smooth(data=qld2, aes(x = NOTIFICATION_DATE, y = n),
                              method='gam')+
geom_point(data=qld2, aes(x = NOTIFICATION_DATE, y = n), alpha=1/4,
             color = 'deeppink4')+ 
theme_update(plot.title = element_text(hjust = 0.5))+
facet_wrap(~AGE_GROUP_5Y, nrow = 4, scales = "free_y")

```



