---
title: "NZ Covid Data Exploration Pt 1"
description: |
  A look at the covid datasets produced by NZ.
author:
  - name: Sean De Boo
    url: https://seandeboo.github.io/hds-blog/
date: 2022-07-17
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
library(tidyverse)
library(plotly)
library(skimr)
library(janitor)
library(plotly)
library(officer)
library(viridis)
library(flextable)

`%notin%` <- Negate(`%in%`)

knitr::opts_chunk$set(echo = FALSE)

```

# NZ Covid Case data exploration

With the recent spike in cases in NZ, and interest in the new Variant BA.5, It would be interesting to look at case numbers by age group.

```{r load, include=FALSE}
## Gather latest case data from NZ
NZ <- read_csv('https://raw.githubusercontent.com/minhealthnz/nz-covid-data/main/cases/covid-case-counts.csv')

## Clean Names
NZ <- clean_names(NZ)

last_date <- max(NZ$report_date)
from_date <- last_date-14
trend_date <- last_date-150

```

### Cases in younger age groups remain low relative to previous peak form the initial opening up wave.

```{r younger, fig.width=14, fig.height=10}
## filter data for 2022 only
NZ2 <- NZ %>% 
  select(-c(case_status, overseas_travel)) %>% 
  filter(report_date >= '2022-03-01') %>%
  filter(age_group != '70+') %>% 
  filter(age_group %notin% c('40 to 49', '50 to 59', '60 to 69','70 to 79', '80 to 89', '90+')) %>% 
  group_by(report_date, age_group) %>% 
  summarise(cases = sum(number_of_cases_reported))

## plot the data for 2022 by age groups

theme_update(plot.title = element_text(hjust = 0.5))
ggplot() +
  ggtitle('NZ confirmed cases') +
  xlab('Date') + ylab('Number of Cases') +
  geom_smooth(data=NZ2, aes(x = report_date, y = cases), method='gam') +
  geom_point(data=NZ2, aes(x = report_date, y = cases, colour=age_group), alpha=1/2)+
  facet_wrap(~age_group, nrow = 3, scale = 'free_y')



```

### Cases in older age groups are climbing fast and above previous peak levels.

```{r older, fig.width=14, fig.height=10}
## filter data for 2022 only
NZ3 <- NZ %>% 
  select(-c(case_status, overseas_travel)) %>% 
  filter(report_date >= '2022-03-01') %>%
  filter(age_group != '70+') %>% 
  filter(age_group %in% c('40 to 49', '50 to 59', '60 to 69','70 to 79', '80 to 89', '90+')) %>% 
  group_by(report_date, age_group) %>% 
  summarise(cases = sum(number_of_cases_reported))

## plot the data for 2022 by age groups

theme_update(plot.title = element_text(hjust = 0.5))
ggplot() +
  ggtitle('NZ confirmed cases') +
  xlab('Date') + ylab('Number of Cases') +
  geom_smooth(data=NZ3, aes(x = report_date, y = cases), method='gam') +
  geom_point(data=NZ3, aes(x = report_date, y = cases, colour=age_group), alpha=1/2)+
  facet_wrap(~age_group, nrow = 3, scale = 'free_y')



```

So far there is no evidence of widespread re-infections

However for 20-29 age group. Reinfections are on the rise. Mostly after a period of 90 days.

``` {r next, fig.width=14, fig.height=10}

  NZ4 <- NZ %>% 
    filter(report_date >= '2022-03-01') %>% 
    select(-c(case_status, overseas_travel, district)) %>% 
    group_by(report_date, age_group, infection_status) %>% 
    filter(infection_status=='Reinfection (< 90 days)') %>% 
    summarise(cases = sum(number_of_cases_reported))
  
  NZ5 <- NZ %>% 
    filter(report_date >= '2022-03-01') %>% 
    select(-c(case_status, overseas_travel, district)) %>% 
    group_by(report_date, age_group, infection_status) %>% 
    filter(infection_status=='Reinfection') %>% 
    summarise(cases = sum(number_of_cases_reported))
  
  ggplot() + 
  geom_point(data=NZ4, aes(x= report_date, y = cases, color= 'Reinfection < 90 days'), alpha=1/2) +
    geom_smooth(data=NZ4, aes(x= report_date, y = cases, method='gam', color= 'Reinfection < 90 days')) +
  geom_point(data=NZ5, aes(x = report_date, y = cases, color= 'Reinfection >90 days'), alpha=1/2) +
    geom_smooth(data=NZ5, aes(x = report_date, y = cases, method='gam', color= 'Reinfection >90 days')) +
  facet_wrap(~age_group, nrow = 3) + 
    labs(x='Date',
         y='Confirmed Cases',
         title='NZ Daily Reinfections by age group')

```

